Intrinsic alignments from projected tidal fields: non-linear impact on cosmic shear probes

### 2024-06-21 sections/1_introduction.tex

Recent cosmic shear measurements from the Kilo Degree Survey\footnote{KiDS:kids.strw.leidenuniv.nl}, the Dark Energy Survey\footnote{DES:www.darkenergysurvey.org}, and the Hyper Suprime Camera Survey\footnote{HSC:www.naoj.org/Projects/HSC} have established weak gravitational lensing as a competitive probe of cosmology \citep[see. \eg][]{KiDS1000_Asgari, KiDS1000_vdB, KiDS1000_Li, DESY3_Secco, DESY3_Amon, HSCY3_Cl, HSCY3_2pcf}, achieving  percent-level measurements of the structure growth parameter $S_8\equiv\sigma_8\sqrt{\Omega_{\rm m}/0.3}$.
The parameters $\Omega_{\rm m}$ and $\sigma_8$, which respectively describe the abundance and the amplitude of the fluctuations in the matter density fluctuations on scales of $8h^{-1}$ Mpc, are highly degenerate in the lensing signal, and additional data (\eg galaxy clustering data) are required to measure the two individually from photometric galaxy surveys \citep{KiDS1000_Heymans, DESY3_3x2, HSCY3_3x2}.
Despite these successful achievements, the current precision of cosmic shear cosmology is mainly limited by the large uncertainty in the intrinsic alignment (IA) of galaxies, a secondary signal that tends to cancel some of the shape correlations produced by lensing \citep[see. \eg][for reviews on IA]{Troxel_IA_review_2015, Kirk_IA_review_2015, Joachimi_IA_review_2015, Kiessling_IA_review_2015,Lamman_IA_guide}. 
If unaccounted for, the IA can bias by 4-5$\sigma$ the inferred cosmological parameters \citep{Kirk2012, Krause2016}.
Additionally, using an inaccurate IA model can substantially impair the inference process, as demonstrated by \citet{DESY3_Secco} in their study on DES-Y3 analysis and by \citet{Paopiamsap2024} in the context of an LSST-like cosmic shear analysis.
%Furthermore, an incorrect IA model can also cause significant damage to the inference, as shown in \citet{DESY3_Secco} in the context of DES-Y3 analysis, or in \citep{Paopiamsap2024} in the context of an LSST-like cosmic shear analysis.

>>>>

Recent cosmic shear measurements from the Kilo Degree Survey\footnote{KiDS:kids.strw.leidenuniv.nl}, the Dark Energy Survey\footnote{DES:www.darkenergysurvey.org}, and the Hyper Suprime Camera Survey\footnote{HSC:www.naoj.org/Projects/HSC} have established weak gravitational lensing as a competitive probe of cosmology \citep[see. \eg][]{KiDS1000_Asgari, KiDS1000_vdB, KiDS1000_Li, DESY3_Secco, DESY3_Amon, HSCY3_Cl, HSCY3_2pcf}, achieving  percent-level measurements of the structure growth parameter $S_8\equiv\sigma_8\sqrt{\Omega_{\rm m}/0.3}$.
The parameters $\Omega_{\rm m}$ and $\sigma_8$, which respectively describe the abundance and the amplitude of the fluctuations in the matter density fluctuations on scales of $8 \: h^{-1}$ Mpc, are highly degenerate in the lensing signal, and additional data (\eg galaxy clustering data) are required to measure the two individually from photometric galaxy surveys \citep{KiDS1000_Heymans, DESY3_3x2, HSCY3_3x2}.
Despite these successful achievements, the current precision of cosmic shear cosmology is mainly limited by the large uncertainty in the intrinsic alignment (IA) of galaxies, a secondary signal that tends to cancel some of the shape correlations produced by lensing \citep[see. \eg][for reviews on IA]{Troxel_IA_review_2015, Kirk_IA_review_2015, Joachimi_IA_review_2015, Kiessling_IA_review_2015,Lamman_IA_guide}. 
If unaccounted for, the IA can bias by 4-5$\sigma$ the inferred cosmological parameters \citep{Kirk2012, Krause2016}.
Additionally, using an inaccurate IA model can substantially impair the inference process, as demonstrated by \citet{DESY3_Secco} in their study on DES-Y3 analysis and by \citet{Paopiamsap2024} in the context of an LSST-like cosmic shear analysis.
%Furthermore, an incorrect IA model can also cause significant damage to the inference, as shown in \citet{DESY3_Secco} in the context of DES-Y3 analysis, or in \citep{Paopiamsap2024} in the context of an LSST-like cosmic shear analysis.
